---
output: html_document
editor_options: 
  chunk_output_type: inline
---

# 1 Exploration

```{r}
test <- read.table(file="test.csv", sep=",", header=TRUE) 
train <- read.table(file="train.csv", sep=",", header=TRUE)

set.seed(100)

```

**Target feature is biodegradability.**

**The target variable is encoded as ready biodegradable (1) and not ready biodegradable (2)**

```{r}
summary(test)
test
```

**How balanced is the target variable?**

Mean of target variable is 1.337 that is how we know there are more not ready biodegradable chemicals in our training dataset than ready biodegradable chemicals.

```{r}
table(train$Class)
```

```{r}
table_class <- (table(train$Class)) 
piepercent <- paste(round(100*table_class/sum(table_class), 2), "%") 
names(table_class) <- c("ready biodegradable", "not ready biodegradable") 
labels <- paste(names(table_class), piepercent) 
pie(table_class, labels = labels, main = "target variable", col=c("pink", "lightblue1")) 
#legend("topright", legend = piepercent, cex = 0.6, fill=c("pink", "lightblue1"))
```

**Are there any missing values present? If there are, choose a strategy that takes this into account.**

```{r}
is.na(train)

train[!complete.cases(train),] 
# 81/1055

#remove instances where one of values is missing 
train <- na.omit(train)
#test <- na.omit(test)
```

**Most of your data is of the numeric type. Can you identify, by adopting exploratory analysis, whether some features are directly related to the target? What about feature pairs?**

```{r}

target_corr = abs(cor(train[,names(train)])[names(train)[42],])
target_corr <- sort(target_corr, decreasing = TRUE)
target_corr_top8 <- target_corr[2: 9]

#the most correlated features
#V1- 0.395
#V27-0.392
#V22-0.364
#V39-0.362
#V15-0.343
#V13-0.342
#V7- 0.332
#V33-0.315

plot(target_corr_top8, xlab="features", ylab = "correlation", main="features with biggest correalton to the target", xaxt = "n")
axis(1, at = c(1,2,3,4,5,6,7,8), labels = names(target_corr_top8))
```



```{r}
correlation_matrix <- cor(train)
correlation_matrix

library(corrplot)
corrplot(correlation_matrix, method = 'color')


```

```{r}
#subset
features = c("V1", "V5", "V7", "V10", "V11", "V13", "V15", "V17", "V18", "V22", "V24", "V27", "V33","V34", "V36", "V38", "V39", "Class")
subset = train[features]
correlation_matrix_small <- cor(subset)
corrplot(correlation_matrix_small, method = 'color')

```

```{r}
pairs(train[c("V1", "V27", "V39", "V12", "V13", "Class")], pch=19, col=c(2, 3))
```

```{r}
library(MASS)
library(ROCR)
linearReg <- lm(V15 ~ V1, data = train)
plot(V15 ~ V1, data = train)
abline(linearReg)
linear <- lda(Class~., train)
```

**Produce at least three types of visualizations of the feature space and be prepared to argue why these visualizations were useful for your subsequent analysis.**

# 2 Modeling


**Try to construct new features from existing ones.**

We looked at attribute information and saw 3 attributes that looked similar:

V5:  F04[C-N]: Frequency of C-N at topological distance 4
V11: F03[C-N]: Frequency of C-N at topological distance 3
V34: F02[C-N]: Frequency of C-N at topological distance 2


We decided to add a new feature that combines these three features and we added the sum of V5, V11 and V34.

```{r}
library(dplyr)
train <- train %>% mutate(V42 = V5 + V11 + V34)
test <- test %>% mutate(V42 = V5 + V11 + V34)
```


First lets split out test set in two sets we will actually use for determening accuracy

```{r}
library(caret)
## Split the data so that we use 70% of it for training
train_index <- createDataPartition(y=train$Class, p=0.7, list=FALSE)

## Subset the data
train_set <- train[train_index, ]
test_set <- train[-train_index, ]
```


***majority classifier***

```{r}

majority.class <- names(which.max(table(train_set$Class)))
majority.class
```

Accuracy of majority classifier:

```{r}
sum(train$Class == majority.class) / length(train$Class)
```

We want to achieve better accuracy than majority classifier.

```{r}
#function for classification accuracy

CA <- function(observed, predicted)
{
  t <- table(observed, predicted)
  
  sum(diag(t)) / sum(t)
}
```

***random classifier***

TODO

***decision tree***

```{r}
library(rpart)
decision_tree <- rpart(Class ~ ., data = train, method = "class")
plot(decision_tree)
text(decision_tree, pretty = 1)
train.dt <- decision_tree
```

```{r}

##########!!!is made with test!!!#######

observed <- test$Class
predicted <- predict(decision_tree, test, type = "class")

confusion_matrix <- table(observed, predicted)
confusion_matrix

CA(observed, predicted)
```

```{r}
library(caret)
importance <- varImp(decision_tree, scale=FALSE)
importance
importance_sorted <- arrange(importance, desc(Overall))
importance_sorted


```


***decision tree with subset of 17 the most important features***

```{r}

features
subset_of_test = test[features]

length(features)

decision_tree_sub <- rpart(Class ~ ., data = subset_of_test, method = "class")
plot(decision_tree_sub)
text(decision_tree_sub, pretty = 1)

observed <- test$Class
predicted <- predict(decision_tree_sub, test, type = "class")

confusion_matrix <- table(observed, predicted)
confusion_matrix

CA(observed, predicted)
```
Accuracy with chosen subset is the same as it was before (with all features)

***Random forest***

```{r}
set.seed(100)
library(randomForest)
train$Class <- as.factor(train$Class)
train.rf <- randomForest(Class ~ ., data = train, trControl=trainControl(method='repeatedcv', number=10, repeats=5, 
                            search = "grid"), metric='Accuracy', ntree = 800, ntry = 30)
train.rf
predicted <- predict(train.rf, test, type="class")
CA(observed, predicted)

```


```{r}
library(dplyr)
var_imp <- varImp(train.rf)
var_imp <- data.frame(variables=row.names(var_imp), importance=var_imp$Overall)

## Create a plot of variable importance
var_imp %>%
        
        ## Sort the data by importance
        arrange(importance) %>%
        
        ## Create a ggplot object for aesthetic
        ggplot(aes(x=reorder(variables, importance), y=importance)) + 
        
        ## Plot the bar graph
        geom_bar(stat='identity') + 
        
        ## Flip the graph to make a horizontal bar plot
        coord_flip() + 
        
        ## Add x-axis label
        xlab('Variables') +
        
        ## Add a title
        labs(title='Random forest variable importance') + 
        
        ## Some layout for the plot
        theme_minimal()
```
```{r}
acc <- c() 

for (i in 1:40){

  feature_importance_sorted <- arrange(var_imp, desc(importance))
  feature_top = feature_importance_sorted[1:i,1]
  feature_top[i+1] = "Class"
  
  library(randomForest)
  set.seed(155)
  train.rf_min <- randomForest(Class ~ ., data = train[feature_top], trControl=trainControl(method='repeatedcv', number=10, repeats=5, 
                              search = "grid"), metric='Accuracy')
  train.rf_min
  predicted <- predict(train.rf_min, test, type="class")
  acc[i] = CA(observed, predicted)
} 
max(acc)
qplot(1:40, acc, xlab = "number of features", ylab = "Performance(acc)", geom = c("point","line"), main="random forest accuracy in relation to number of features")

```  

***KNN***


```{r}
library(caret)
library(CORElearn)
library(ggplot2)


train$Class = factor(train$Class)

grid = expand.grid(k = c(1:20))

knn_model = train(Class ~., method= "knn", data = train, 
                    trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3, search = "grid"),
                     tuneGrid = grid, preProcess = c("scale", "center"))

knn_model
best_k = which.max(knn_model$results$Accuracy)

qplot(1:20, knn_model$results$Accuracy, xlab = "k", ylab = "Performance(acc)", geom = c("point","line"))

train.knn <- knn_model

```

On the graph we can see the best K for KNN in interval 1-20 is 5

```{r}
### Made with test !!!! ###

train.lda <- lda(Class~., train)
train.lda.values <- predict(train.lda, train)
ldahist(train.lda.values$x[,1], g=train$Class)

train.lda.acc <- CA(test$Class, predict(train.lda, test)$class)

train.lda.conf <- table(test$Class, predict(train.lda, test)$class)

```
