---
output: html_document
editor_options: 
  chunk_output_type: inline
---

# 1 Exploration

```{r}
test <- read.table(file="test.csv", sep=",", header=TRUE) 
train <- read.table(file="train.csv", sep=",", header=TRUE)
dataset = rbind(test, train)

```

**Target feature is biodegradability.**

**The target variable is encoded as ready biodegradable (1) and not ready biodegradable (2)**

```{r}
summary(dataset)
dataset
```

**How balanced is the target variable?**

Mean of target variable is 1.337 that is how we know there are more not ready biodegradable chemicals in our training dataset than ready biodegradable chemicals.

```{r}
table(dataset$Class)
table(test$Class)
table(train$Class)
```

```{r}
table_class <- (table(dataset$Class)) 
piepercent <- paste(round(100*table_class/sum(table_class), 2), "%") 
names(table_class) <- c("ready biodegradable", "not ready biodegradable") 
labels <- paste(names(table_class), piepercent) 
pie(table_class, labels = labels, main = "target variable", col=c("pink", "lightblue1")) 
#legend("topright", legend = piepercent, cex = 0.6, fill=c("pink", "lightblue1"))
```

**Are there any missing values present? If there are, choose a strategy that takes this into account.**

```{r}
is.na(dataset)

dataset[!complete.cases(dataset),] 
# 81/1055

#remove instances where one of values is missing 
train <- na.omit(train)
test <- na.omit(test)
dataset <- na.omit(dataset)
dataset
```

```{r}
table_class <- (table(dataset$Class)) 
piepercent <- paste(round(100*table_class/sum(table_class), 2), "%") 
names(table_class) <- c("ready biodegradable", "not ready biodegradable") 
labels <- paste(names(table_class), piepercent) 
pie(table_class, labels = labels, main = "target variable (without instances with missing values)", col=c("pink", "lightblue1")) 

```

**Most of your data is of the numeric type. Can you identify, by adopting exploratory analysis, whether some features are directly related to the target? What about feature pairs?**


```{r}

target_corr = abs(cor(dataset[,names(dataset)])[names(dataset)[42],])
target_corr

#the most correlated features
#V1- 0.395
#V27-0.392
#V22-0.364
#V39-0.362
#V15-0.343
#V13-0.342
#V7- 0.332
#V33-0.315


```

TODO for pairs

**Produce at least three types of visualizations of the feature space and be prepared to argue why these visualizations were useful for your subsequent analysis.**



# 2 Modeling

***majority classifier***

```{r}
majority.class <- names(which.max(table(train$Class)))
majority.class
```

Accuracy of majority classifier:

```{r}
sum(test$Class == majority.class) / length(test$Class)
```

We want to achieve better accuracy than majority classifier.

```{r}
#function for classification accuracy

CA <- function(observed, predicted)
{
  t <- table(observed, predicted)
  
  sum(diag(t)) / sum(t)
}
```

***random classifier***

TODO

***decision tree***

```{r}
# . pomeni da pogledamo za vse atribute
library(rpart)
?rpart
dt <- rpart(Class ~ ., data = test, method = "class")
plot(dt)
text(dt, pretty = 1)
```

```{r}
observed <- test$Class
predicted <- predict(dt, test, type = "class")

confusion_matrix <- table(observed, predicted)
confusion_matrix

CA(observed, predicted)
```

Decstion tree classifies better than majority classifier


