---
output: html_document
editor_options: 
  chunk_output_type: inline
---

# 1 Exploration

```{r}
test <- read.table(file="test.csv", sep=",", header=TRUE) 
train <- read.table(file="train.csv", sep=",", header=TRUE)
dataset = rbind(test, train)

```

**Target feature is biodegradability.**

**The target variable is encoded as ready biodegradable (1) and not ready biodegradable (2)**

```{r}
summary(dataset)
dataset
```

**How balanced is the target variable?**

Mean of target variable is 1.337 that is how we know there are more not ready biodegradable chemicals in our training dataset than ready biodegradable chemicals.

```{r}
table(dataset$Class)
table(test$Class)
table(train$Class)
```

```{r}
table_class <- (table(dataset$Class)) 
piepercent <- paste(round(100*table_class/sum(table_class), 2), "%") 
names(table_class) <- c("ready biodegradable", "not ready biodegradable") 
labels <- paste(names(table_class), piepercent) 
pie(table_class, labels = labels, main = "target variable", col=c("pink", "lightblue1")) 
#legend("topright", legend = piepercent, cex = 0.6, fill=c("pink", "lightblue1"))
```

**Are there any missing values present? If there are, choose a strategy that takes this into account.**

```{r}
is.na(dataset)

dataset[!complete.cases(dataset),] 
# 81/1055

#remove instances where one of values is missing 
train <- na.omit(train)
test <- na.omit(test)
dataset <- na.omit(dataset)
dataset
```

```{r}
table_class <- (table(dataset$Class)) 
piepercent <- paste(round(100*table_class/sum(table_class), 2), "%") 
names(table_class) <- c("ready biodegradable", "not ready biodegradable") 
labels <- paste(names(table_class), piepercent) 
pie(table_class, labels = labels, main = "target variable (without instances with missing values)", col=c("pink", "lightblue1")) 

```

**Most of your data is of the numeric type. Can you identify, by adopting exploratory analysis, whether some features are directly related to the target? What about feature pairs?**

```{r}

target_corr = abs(cor(dataset[,names(dataset)])[names(dataset)[42],])
target_corr <- sort(target_corr, decreasing = TRUE)
target_corr_top8 <- target_corr[2: 9]

#the most correlated features
#V1- 0.395
#V27-0.392
#V22-0.364
#V39-0.362
#V15-0.343
#V13-0.342
#V7- 0.332
#V33-0.315

plot(target_corr_top8, xlab="features", ylab = "correlation", main="features with biggest correalton to the target", xaxt = "n")
axis(1, at = c(1,2,3,4,5,6,7,8), labels = names(target_corr_top8))
```


```{r}
correlation_matrix <- cor(dataset)
correlation_matrix

library(corrplot)
corrplot(correlation_matrix, method = 'color')


```
```{r}
#subset
features = c("V1", "V5", "V7", "V10", "V11", "V13", "V15", "V17", "V18", "V22", "V24", "V27", "V33", "V36", "V38", "V39", "Class")
subset = dataset[features]
print(subset)
correlation_matrix_small <- cor(subset)
corrplot(correlation_matrix_small, method = 'color')

```

```{r}

linearReg <- lm(V15 ~ V1, data = dataset)
plot(V15 ~ V1, data = dataset)
abline(linearReg)

```

**Produce at least three types of visualizations of the feature space and be prepared to argue why these visualizations were useful for your subsequent analysis.**

# 2 Modeling

***majority classifier***

```{r}
majority.class <- names(which.max(table(train$Class)))
majority.class
```

Accuracy of majority classifier:

```{r}
sum(test$Class == majority.class) / length(test$Class)
```

We want to achieve better accuracy than majority classifier.

```{r}
#function for classification accuracy

CA <- function(observed, predicted)
{
  t <- table(observed, predicted)
  
  sum(diag(t)) / sum(t)
}
```

***random classifier***

TODO

***decision tree***

```{r}
# . pomeni da pogledamo za vse atribute
library(rpart)
decision_tree <- rpart(Class ~ ., data = test, method = "class")
plot(decision_tree)
text(decision_tree, pretty = 1)
```

```{r}
observed <- test$Class
predicted <- predict(decision_tree, test, type = "class")

confusion_matrix <- table(observed, predicted)
confusion_matrix

CA(observed, predicted)
```
```{r}
library(caret)
importance <- varImp(decision_tree, scale=FALSE)
importance_sorted <- sort(t(importance), decreasing = TRUE, index.return = TRUE)
print(t(importance))
importance_sorted
# summarize importance
print(importance)

#V15
#V1
#V17
#V10
#V12
#V13
#V11
#V16

```

Decision tree classifies better than majority classifier
***decision tree with subset***
```{r}

features
subset_of_test = test[features]

decision_tree_sub <- rpart(Class ~ ., data = subset_of_test, method = "class")
plot(decision_tree_sub)
text(decision_tree_sub, pretty = 1)

observed <- test$Class
predicted <- predict(decision_tree_sub, test, type = "class")

confusion_matrix <- table(observed, predicted)
confusion_matrix

CA(observed, predicted)
```
Accuracy with choosen subset is the same as it was before (with all features)

***KNN***

```{r}

library(CORElearn)
library(ggplot2)

foldIdx <- cvGen(nrow(train), k=3)
evalCore<-list()
overallScores <- c()

for (x in 1:20) {
  
  for (j in 1:3){
    dTrain <- test[foldIdx!=j,]
    dTest  <- test[foldIdx==j,]
    dTest <- na.omit(dTest)
    dTrain <- na.omit(dTrain)
    model_knn <- CoreModel(Class ~ ., data = train, model="knn", kInNN = x)
    
    predCore <- predict(model_knn, dTest)
  
    
    evalCore[[j]] <- modelEval(model_knn, correctClass = dTest$Class, predictedClass = predCore$class,   
                                 predictedProb=predCore$prob)
    
    #evalCore[[j]] <- modelEval(modelCore, correctClass = dTest$Class, predictedClass = predCore$class, 
    #                            predictedProb=predCore$prob )
    destroyModels(model_knn)
  }
  
  results <- gatherFromList(evalCore)
  meanPerformances <- sapply(results, mean)
  overallScores <- c(overallScores, meanPerformances["accuracy"])
}

bestK <- which.max(overallScores)

cm.knn <- CoreModel(Class ~ ., data = train, model="knn", kInNN = bestK)

predicted_knn <- predict(cm.knn, test, type="class")
  
accuracy = CA(observed, predicted_knn)
accuracy

qplot(1:20, overallScores, xlab = "k", ylab = "Porvormance(acc)", geom = c("point","line"))

#bestK:    9
#accuracy: 0.8421053

#8  -> 0.8277512
#9  -> 0.8421053
#10 -> 0.8229665
#15 -> 0.8086124
#17 -> 0.8229665
#20 -> 0.8133971


```






